---
author: "Aida Aquino Arias"
---
For this lab, you will be joining and filtering related data sets to solve a murder mystery!

# Part 1: GitHub Workflow

At the top of the document (in the YAML) there is an `author` line that says 
`"Your name here!"`. Change this to be your name and save your file either by
clicking on the blue floppy disk or with a shortcut (command / control + s). 

Be sure to [commit the files to your repo](https://happygitwithr.com/existing-github-first#stage-and-commit). 

Let's get started!


# Part 2: Some Words of Advice

-   Set chunk options carefully.

-   Make sure you don't print out more output than you need.

-   Make sure you don't assign more objects than necessary---avoid "object junk"
in your environment.

-   Make your code readable and nicely formatted.

-   Think through your desired result **before** writing any code.

# Part 3: Finding the Killer

Northwestern University's Knight Lab wanted to help sharpen users' database
skills, so they created a murder mystery. Can you solve this crime in SQL City??

The relational data you will be working with contains tables with different
pieces of information pertinent to the crime - people, social media check-ins,
driver's licenses, crime scene reports, police interviews, and more!


## Access the Data

This code chunk will read in **all** of the tables of data for you. Don't modify or
remove this! Take some time to look at each file type so that 

```{r}
#| message: false
#| warning: false
#| label: read-in-data

library(tidyverse)

# If purrr is not detected, install the package
if (!"purrr" %in% installed.packages()) install.packages("purrr")

source("https://raw.githubusercontent.com/jcanner/stat_210_2025_website/main/labs/instructions/lab-7-setup.R")
```

## Solve the Crime

### Crime Scene Report

Detective Wickham reaches out to you...

> A crime has taken place and I need your help! There was a murder in SQL City sometime on January 15, 2018. Could you retrieve the crime scene report from the police department’s database and follow the clues to find the person responsible for the murder?!


**Step 1: Find the police report in `crime_scene_report`. Then used the information in the report to move on to the next data source based on the information learned.**

```{r}
# Code for looking at the relevant crime scene report.

#created data file to store in the filtered crime report for the murder and city
report_cleaned <- crime_scene_report |>
  filter(city == "SQL City", type == "murder")

# created data file to contain the data of witness one by where they live, using arrange, desc and slice to be able to find the last house on the street
witness_one <- person |>
  filter(address_street_name %in% c("Northwestern Dr")) |>
  arrange(desc(address_number)) |>
  slice_tail(n=1)

#created data file to contain the data of witness two, by filtering the street name and first name. First separating the name into first and last to be able to locate the first name only.
witness_two <- person |>
  separate(name, into = c("first_name", "last_name"), sep = " ")|>
  filter(address_street_name == "Franklin Ave", first_name == "Annabel")

#joined the witness' into one data file, people_data, then joined the interview to see if any of the witness gave a testimony. By comparing the witness id and the interviewee testimony, Annabel gave a recount of the murder.
people_data <- full_join(witness_one, witness_two)
people_data <- inner_join(people_data, interview, by = c("id" = "person_id"))
#finding correlation between the gym members


#created a data file to keep the info for the potential suspect, narrowing it down by first joining the member and check in data files to find any correlation and the full info such as name, and gym start date. Then changing the format of the dates so that it can be easy to read.
gym_member <- full_join(get_fit_now_check_in, get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(check_in_date = ymd(check_in_date)) |>
  mutate(year = year(check_in_date),
         month = month(check_in_date), 
         day = day(check_in_date)) |>
#filtering out the date that was given by Annabel and then once again making the start date easy to read by mutating it. Finally filtering to include only people that started their gym membership before Jan. 9th and people Annabel could not have seen since they checked out before she arrived.  
  filter(year == 2018, month == 1, day == 9) |>
  mutate(membership_start_date = ymd(membership_start_date)) |>
  filter(membership_start_date <= "2018-01-09", check_out_time >= 1600) |>
  left_join(interview)
# since I got my suspect I decided to see who attended the SQL Symphony Concert 3 times. First separating the date to make it easier to search by only year and month, after filtering out the even by the date I then sorted the numbers to make it easier to see who attended at least 3 times. Once i narrowed it down I joined the person data to be able to find the name of the suspect. Then I goind the drivers lices data in order to correctly identify Miranda as the one who hired Jeremy.
suspect <- facebook_event_checkin|>
  mutate(date = ymd(date)) |>
  mutate(year = year(date),
         month = month(date), 
         day = day(date)) |>
  filter(year == 2017, month == 12 & event_name == "SQL Symphony Concert") |>
  count(person_id, sort = TRUE) |>
  right_join(person, by = c("person_id" = "id")) |>
  filter(name == "Miranda Priestly") |>
  left_join(drivers_license, by = c("license_id" = "id"))
print(suspect)
```




**Next Steps: Follow the evidence to the person responsible for the murder, building a report as you go.**
There are accomplices, some knowingly and some unknowingly, but there is only one mastermind.


Solve the murder mystery, showing **all of your work in this document**. Your
document and code must be well organized, easy to follow, and reproducible.

+ Use headers and written descriptions to indicate what you are doing.
+ Use `dplyr` verbs and `join` functions rather than just looking through the
tables manually. Functions from `stringr` and `lubridate` will also be useful.
+ Use good code formatting practices.
+ Comment your code.
+ Cite any external sources you use to solve the mystery.



::: callout-note
## And the final suspect is...

*put the name of the person responsible for the murder here: Miranda Priestly ordered the murder and Jeremy Bowers executed the murder.*
:::
